---
title: "RRWM Data Activity - Workflow"
author: "Julia"
date: "21/09/2021"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dataset

Dataset being used: Canadian Community Health Survey (CCHS) - Mental Health, 2012

Link to datset: https://abacus.library.ubc.ca/dataset.xhtml?persistentId=hdl:11272.1/AB2/2PSQXJ

Dataset citation: Statistics Canada. Health Statistics Division, 2014, "Canadian Community Health Survey (CCHS) - Mental Health, 2012", https://hdl.handle.net/11272.1/AB2/2PSQXJ, Abacus Data Network, V1, UNF:6:NHaAsIbeTwjPzakM8IWR9w== [fileUNF]

# Import the data
- Install package: "haven"
- Import the dataset into R
```{r, echo = FALSE, warning = FALSE}
# install.packages("haven")
library(haven)
cchs2012original <- read_sav("cchs2012_mh.sav")
cchs2012raw <- zap_labels(cchs2012original)
```

## Data cleaning

- Install packages for data cleaning: "psych", "readr", "dplyr"
```{r, echo = FALSE, results = 'hide', warning = FALSE, error = FALSE, message = FALSE}
# install.packages("psych")
library(psych)
# install.packages("readr")
library(readr)  
# install.packages("dplyr")
library(dplyr)
```

## Extract raw variables

- Age: DHHGAGE
- Sex: DHH_SEX
- Marital status: DHHGMS
- Education: EDUDR04
- Negative social interactions: NSIDSC
- General life satisfaction: GENGSWL

```{r, echo = FALSE, results = 'hide', error = FALSE, message = FALSE}
# Extract variables needed ------------------------------------------------

cchs2012 <- 
  cchs2012raw %>% select(DHHGAGE,DHH_SEX,DHHGMS,EDUDR04,NSIDSC,GENGSWL)
```

## Variable transformations:

- Age: Factor variable (from 0 = age group 15-19 to 13 = age 80+)
- Sex: Factor variable (0 = male, 1 = female)
- Marital status: Factor variable (0 = married, 1 = common-law, 2 = widowed, 3 = divorced/separated, 4 = single)
- Education: Factor variable (0 = < post secondary, 1 = secondary school, 2 = some post-secondary, 3 = post-secondary)
- Social strain: Continuous variable (0-12 scale)
- Satisfaction with life: Continuous variable (5 = Very satisfied, 4 = Satisfied, 3 = Neither satisfied or dissatisfied, 2 = Dissatisfied, 1 = Very dissatisfied)

```{r, echo = FALSE, results = 'hide', error = FALSE, message = FALSE}
# Age ---------------------------------------------------------------------

table(cchs2012$DHHGAGE, useNA='always')
# Original coding: 
# 1 = 'Age 15-19'
# 2 = 'Age 20-24'
# 3 = 'Age 25-29'
# 4 = 'Age 30-34'
# 5 = 'Age 35-39'
# 6 = 'Age 40-44'
# 7 = 'Age 45-49'
# 8 = 'Age 50-54'
# 9 = 'Age 55-59'
# 10 = 'Age 60-64'
# 11 = 'Age 65-69'
# 12 = 'Age 70-74'
# 13 = 'Age 75-79'
# 14 = 'Age 80+'
cchs2012 <- cchs2012 %>% rename(age = DHHGAGE) #Renaming variable to age
cchs2012 <- cchs2012 %>% 
  mutate(age = dplyr::recode(cchs2012$age, '1'=0,'2'=1, '3'=2,'4'=3, 
                             '5'=4,'6'=5, '7'=6,'8'=7, '9'=8,'10'=9, '11'=10,
                             '12'=11, '13'=12,'14'=13)) # Recoding 0 as reference group
cchs2012$age_f <- 
  factor(cchs2012$age, levels=c(0,1,2,3,4,5,6,7,8,9,10,11,12,13), 
         labels=c('Age 15-19','Age 20-24', 'Age 25-29', 'Age 30-34',
                  'Age 35-39', 'Age 40-44', 'Age 45-49', 'Age 50-54', 
                  'Age 55-59','Age 60-64', 'Age 65-69', 'Age 70-74', 
                  'Age 75-79', 'Age 80+'))

# Sex ---------------------------------------------------------------------

table(cchs2012$DHH_SEX, useNA='always') 
# Original coding: 
# 1 = male
# 2 = female
cchs2012 <- cchs2012 %>% rename(sex = DHH_SEX) # Renaming variable to sex
cchs2012 <- cchs2012 %>% 
  mutate(sex = dplyr::recode(cchs2012$sex, '1'=0,'2'=1)) 
# Recoding 0=male, 1=female

cchs2012$sex_f <- 
  factor(cchs2012$sex, levels=c(0,1), labels=c('male','female'))


# Marital status ----------------------------------------------------------

table(cchs2012$DHHGMS, useNA='always')
# Original coding: 
# 1 = married
# 2 = common-law
# 3 = widowed
# 4 = divorced/separated
# 5 = single
cchs2012 <- cchs2012 %>% rename(marst = DHHGMS)
cchs2012 <- cchs2012 %>% 
  mutate(marst = dplyr::recode(cchs2012$marst, '1'=0,'2'=1, '3'=2,'4'=3, 
                             '5'=4)) # Recoding 0 as reference group
cchs2012$marst_f <- 
  factor(cchs2012$marst, levels=c(0,1,2,3,4), labels=c('married', 'common-law',
                                                       'widowed', 'divorced/separated',
                                                       'single'))


# Education ---------------------------------------------------------------

table(cchs2012$EDUDR04, useNA='always')
# Original coding: 
# 1 = Less than secondary school graduation
# 2 = Secondary school graduation, no post-secondary
# 3 = Some post-secondary education
# 4 = Post-secondary certificate/diploma or university degree
cchs2012 <- cchs2012 %>% rename(education = EDUDR04)
cchs2012 <- cchs2012 %>% 
  mutate(education = dplyr::recode(cchs2012$education, '1'=0,'2'=1, '3'=2,'4'=3)) 
# Recoding 0 as reference group
cchs2012$education_f <- 
  factor(cchs2012$education, levels=c(0,1,2,3), 
         labels=c('< post secondary', 'secondary school', 
                  'some post-secondary', 'post-secondary'))


# Negative social strain --------------------------------------------------
table(cchs2012$NSIDSC, useNA='always')
# Original coding: 0-12 scale
cchs2012 <- cchs2012 %>% rename(socialstrain = NSIDSC)
# hist(cchs2012$socialstrain)


# Satisfaction with Life --------------------------------------------------

table(cchs2012$GENGSWL,useNA = 'always') # reverse
# Original coding:
# 1 = Very satisfied
# 2 = Satisfied 
# 3 = Neither satisfied or dissatisfied
# 4 = Dissatisfied
# 5 = Very dissatisfied
cchs2012 <- cchs2012 %>% rename(swl = GENGSWL)
cchs2012 <- cchs2012 %>% 
  mutate(swl = dplyr::recode(cchs2012$swl, '1'=5,'2'=4,'3'=3,'4'=2, '5'=1))
# New coding: 
# 5 = Very satisfied
# 4 = Satisfied 
# 3 = Neither satisfied or dissatisfied
# 2 = Dissatisfied
# 1 = Very dissatisfied
# hist(cchs2012$swl) #histogram
```


## Define Analytic Sample

- Exclude participants in any age group aged <60.

```{r, echo = FALSE, results = 'hide', error = FALSE, message = FALSE}

# Excluding participants --------------------------------------------------

#excluding anyone under the age of 60
cchs2012final <- subset(cchs2012, age>8)
```


## Analyses

### Analysis 1: Regression line

- Create a graph of the unadjusted regression with social strain (outcome variable) and satisfaction with life (exposure variable).

```{r, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE}

# Plotting negative social interactions and life satisfaction
library(ggplot2)
g1 <- ggplot(cchs2012final, mapping = aes(x = socialstrain, y = swl))
g1 +
  geom_smooth(method = "lm") # use linear regression model

```


### Analysis 2: Linear regression

- Conduct a linear regression where the outcome is satisfaction with life and the exposure is negative social interactions.
- Conduct a second linear regression where the outcome is satisfaction with life, and the exposures are: negative social interactions, age, sex, marital status, and education.
- Place these side-by-side in the same table.
- Install package: "sjPlot"
```{r, echo = FALSE, results = FALSE, warning = FALSE}

# Model 1: Negative social interactions & satisfaction with life, unadjusted
mod1 <- lm(swl ~ 1 + socialstrain, data=cchs2012)
summary(mod1)

# Model 2: Negative social interactions & satisfaction with life, adjusted

mod2 <- lm(swl ~ 1 + socialstrain + age_f + sex_f + marst_f + education_f,
           data=cchs2012final)
summary(mod2)

```
```{r, echo = FALSE, message = FALSE}
# install.packages("sjPlot")
library(sjPlot)
tab_model(mod1,mod2)
```

