---
title: "RRWM Data Activity - Replicating Sofia's Code"
author: "Julia"
date: "21/09/2021"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# See notes/preliminary reflections at the bottom of this document.

# Dataset

Dataset being used: Canadian Community Health Survey (CCHS) - Mental Health, 2012

Link to datset: https://abacus.library.ubc.ca/dataset.xhtml?persistentId=hdl:11272.1/AB2/2PSQXJ

Dataset citation: Statistics Canada. Health Statistics Division, 2014, "Canadian Community Health Survey (CCHS) - Mental Health, 2012", https://hdl.handle.net/11272.1/AB2/2PSQXJ, Abacus Data Network, V1, UNF:6:NHaAsIbeTwjPzakM8IWR9w== [fileUNF]

# Import the data
- Install package: "haven"
- Import the dataset into R
```{r, echo = FALSE, warning = FALSE}
# install.packages("haven")
library(haven)
library(tidyverse)
cchs2012original <- read_sav("cchs2012_mh.sav")
cchs2012 <- zap_labels(cchs2012original)
```

## Excluding Participants

From the variable “PHSGAPA” (Moderate/vigorous physical activity -average/hours) drop those
individuals whose content is “NOT STATED”.

```{r}
# The variable of interest is "PHSGAPA" 99.99 means "NOT STATED"
cchs2012<-cchs2012%>%filter(PHSGAPA<99)
```

## Transforming variables

### Transform the variable sex (DHH_SEX) into factor

```{r}
# Sex ---------------------------------------------------------------------
cchs2012$DHH_SEX<-factor(cchs2012$DHH_SEX,
                         levels = c(1,2),
                         labels = c("MALE","FEMALE"))
```

### Transform the physical activity variable

Create the variable “GROUP” by transforming “PHSGAPA” into the next categories:
– “NOT” if they do 0 hours of exercise
– “0%-25%” for those in the first quantile without considering those that do 0 hours of exercise.
– “25-50” for those in the second quantile without considering those that do 0 hours of exercise.
– “50-75” for those in the third quantile without considering those that do 0 hours of exercise.
– “75-100” for those in the fourth quantile without considering those that do 0 hours of exercise.

```{r}

# Those that do not exercise = "NOT"
noex <- cchs2012 %>%
  filter(PHSGAPA==0)

noex$GROUPS <- "NOT"
# Those that DO exercise
doex<-cchs2012 %>% 
  filter(PHSGAPA>0)

doex$GROUPS <- with(doex, cut(PHSGAPA,
breaks = quant <- quantile(PHSGAPA),
labels = paste0(names(quant)[-5],"-",names(quant)[-1]),
include.lowest=TRUE))

# Combine
COMBINED <- rbind(noex,doex)

```

# Table

Create a table that shows the percentage of population in each “GROUP” by age-groups breaking down
by sex.

```{r}
A1 <- seq(15,80,5)
A2 <- A1+4; A2[length(A2)]="more"
Alabels <- paste0(A1,"-",A2)

COMBINED$DHHGAGE <- factor(COMBINED$DHHGAGE,
levels = 1:14,
labels = Alabels)
TABLE=COMBINED %>%
  group_by(GROUPS,DHHGAGE,DHH_SEX)%>%
  count(DHHGAGE,DHH_SEX)%>%
  ungroup()%>%
  group_by(GROUPS,DHH_SEX)%>%
  mutate(perc=100*n/sum(n))
```

```{r}
(TABLE %>%
   ggplot() +
   geom_tile(aes(GROUPS,DHHGAGE,fill= perc))+
   geom_text(aes(GROUPS,DHHGAGE,label = round(perc, 1)),size=2.8) +
   facet_grid(DHH_SEX~.)+
   theme_minimal(base_size = 8)+
   theme(axis.text.x = element_text(hjust = 0.5,size = 8),
         axis.title = element_blank(),
         legend.position= "right",
         plot.caption = element_text(hjust =1,size = 8))+
   scale_fill_distiller(palette = "Spectral",trans = 'log10')+
   guides(fill=guide_colorbar(title = "%",
                              title.position = "right",
                              title.hjust = 0.5,
                              barwidth = 0.5, barheight = 10,
                              ticks.colour = "black",
                              ticks.linewidth = 8,ticks = TRUE)))
```

Transform to NA all the values in “GEN_02A2”,“GEN_04”,“GEN_07”, and “GEN_10” that correspond
to responds: “DON’T KNOW”, “REFUSAL”, and “NOT STATED”.

```{r}
VARIABLES<-c("GEN_02A2","GEN_04","GEN_07","GEN_10")

cchs2012b <- cchs2012%>%
  select(c("PHSGAPA","DHH_SEX","DHHGAGE", "GEN_02A2","GEN_04",
           "GEN_07","GEN_10","GENDHDI"))

# GEN_02A2: Satisfaction with life in general - 11 point scale GEN 11
cchs2012b<-cchs2012b%>%mutate(GEN_02A2=ifelse(GEN_02A2<97,GEN_02A2,NA))

# GEN_04: Frequency - trouble sleeping
cchs2012b<-cchs2012b%>%mutate(GEN_04=ifelse(GEN_04<7,GEN_04,NA))

# GEN_07: Perceived life stress
cchs2012b<-cchs2012b%>%mutate(GEN_07=ifelse(GEN_07<7,GEN_07,NA))

# GEN_10: Sense of belonging to local community 
cchs2012b<-cchs2012b%>%mutate(GEN_10=ifelse(GEN_10<7,GEN_10,NA))

# GENDHDI: Perceived health
cchs2012b<-cchs2012b%>%mutate(GENDHDI=ifelse(GENDHDI<9,GENDHDI,NA))

```



# Linear Regression Table

```{r}
# Run a linear model where the outcome is “PHSGAPA” and the independent variables are “DHH_SEX”,“DHHGAGE”, “GEN_02A2”,“GEN_04”,“GEN_07”,“GEN_10”, and “GENDHDI”.

fit <- lm(PHSGAPA~., data = cchs2012b)
3
SUM <- summary(fit)

```
Report the results as a table (Table 1) with the coefficients and their 95% confidence intervals, standard
errors, p-values, and the standard p-value< ’*’ coding. Also rename the variables into more descriptive
names:
– DHH_SEXFEMALE: “Sex:Female”
– DHHGAGE: “Age”
– GEN_02A2: “Satif. Life”
– GEN_04: “Trob. Sleep.”
– GEN_07: “Life Stress”
– GEN_10: “Sense Belong.”
– GENDHDI: “Health”

```{r}
# install.packages("kableExtra")

library(kableExtra)

Coef <- coef(fit)
# Confidence Interval
conf<-confint(fit)
CI1<-data.frame(lower=round(conf[,1],4),
mean=round(Coef,4),
upper=round(conf[,2],4))

# SE
CI1$SE <- SUM$coefficients[,2]

# p-values
p <- SUM$coefficients[,4]
CI1$`p-value`=round(p,digits = 4)
CI1$NAMES=row.names(CI1)
CI<-rbind(CI1)
CI$`p-value<`<-c("***","**","*","."," ")[findInterval(CI$`p-value` ,
sort(c(1, 0.1, 0.05, 0.01, 0.001, 0))) ]
CI<-CI[-1,]
CI$NAMES <- factor(CI$NAMES,
levels = rev(CI$NAMES),
labels = rev(c("Sex:Female","Age","Satif. Life","Trob. Sleep.",
"Life Stress","Sense Belong.","Health")))
CI<-CI[,c(6,1:5,7)]
row.names(CI) <- NULL

kableExtra::kbl(CI, booktabs = T, caption="Results Liner Model") #%>%
# kableExtra::kable_styling(position = "center")
```

# Plot the results
```{r}
library(ggplot2)

ggplot(CI, aes(x=NAMES, y=mean,label=`p-value<`)) +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=.5,position = position_dodge(1)) +
  geom_point(size=2,position = position_dodge(1))+
  coord_flip()+
  scale_x_discrete(labels = levels(CI$NAMES), breaks=levels(CI$NAMES),
                   expand = c(0.1,0.1))+
  geom_hline(yintercept = 0, linetype="dashed",color="red")+
  theme(panel.spacing = unit(1, "lines"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "top",text = element_text(size = 10),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.y = element_line(size = 0.25,
                                          linetype = 'dashed',
                                          colour = "gray"),
        panel.grid.major.x = element_line(size = 0.25,
                                          linetype = 'dashed',
                                          colour = "gray"),
        strip.background = element_rect(color = NA,
                                        fill = NA, size = 1),
        axis.line = element_line(color = 'black'),
        plot.caption = element_text(hjust = 0.5,size = 10))+
  geom_text(hjust=0, vjust=0,size=8,position = position_dodge(1))
```

# Notes:

- General lack of knowledge about R --> reproducibility problems (on my end)
- General lack of knowledge about functions to be used --> reproducibility problems (on my end)
- Did most of the data cleaning and variable tranformations on my own, had to use Sofia's code to make the figures 
- Would have been nice to know what packages in R to use to reproduce graphs
- Transform the variable sex (DHH_SEX) into factor - Sex variable: Didn't know it should have been kept as 1 & 2, 
(I changed it to 0 and 1 because that's what I normally do.)
- Didn't have the color palette needed to recreate first table

